\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-07-03 forwards-and-backwards.tex
.

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd

@@

@@

@@

% % % % % % % % % % % % % %
\subsection{Structure of the paper}

In the paper we study scaling limits of spatial populations,
obtaining convergence of both the population process
(i.e., the population density as a function of time,
although strictly speaking it is a measure that may not have a density)
as well as of lineages traced back through such a population.
We retain linages through the rescaling process
by means of a lookdown construction.
First, in section~\ref{sec: Model and main results},
we describe the model and the main theorems
(Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}).
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results,
namely, X Y and Z \comment{TODO}.
In section \ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them.
In Section \ref{sec:lookdown} we define and discuss the lookdown process that retains the levels.
This is followed by the technical proofs in section \ref{sec:proofs}.

@@

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are bounded, uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR_{\ge 0}$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
and that $(x, m) \mapsto m^2 \gamma(x, m)$ is uniformly bounded.
Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F: \IR^d \times \IR_{\ge 0} \to \IR$ is bounded and uniformly continuous
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    r(x, \smooth{r} \eta(x))
    \gamma(\smooth{\gamma} \eta(x))
    - \frac{1}{\theta}
    F(x, \smooth{F} \eta(x)) .
\end{align}
\end{definition}

@@

% % % % % % % % % % % % % % %
\subsection{Main results on population processes}

Our main results depend on two dichotomies:
Is the limiting process deterministic or a (generalized) superprocess?
And, are interactions local in the limit or not?
Below we have results for deterministic limits with local and nonlocal interactions,
and for superprocess limits with nonlocal interactions.
See Figure \ref{fig:super_vs_det_2d} for illustrative simulations.
We suspect that limits are possible the remaining (local, superprocess) case
in only $d=1$, but do not discuss the case further.

@@

    The sequence of stochastic processes
    $(\eta^N_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f:\mathbb{R}^d \rightarrow \mathbb{R}$,
    \begin{align} \label{eqn:mgale_problem}
        M_t
        &:=
        \langle f(x), \eta_t(dx) \rangle
        -
        \langle f(x), \eta_0(dx) \rangle
        \\ & \qquad
        -
        \int_0^t \big\langle
            \gamma(x, \smooth{\gamma} \eta_{s}(x))
            \mathcal{B}\left(
                f(\cdot) r(\cdot, \smooth{r} \eta_{s}(\cdot))
            \right)(x)
            +
            f(x)
            F(x, \smooth{F} \eta_{s}(x)),
            \eta_{s}(dx)
        \big\rangle ds,
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x,  \eta_{s} \right)
            r\left(x, \eta_{s} \right)+\mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

@@

% % % % % % % % % % % %
\subsection{Main results on lineages}

@@

% % % % % % % % % % % %
\subsection{A nonlinear diffusive population: the Porous Medium Equation}

@@

% % % % % % % % % % % %
\subsection{Traveling waves}

@@

% % % % % % % % % % % % % % % %
\subsection{Nonuniqueness of population density}

@@

% % % % % % % % % % % % % % % % % %
\subsection{Population processes}
    \label{sec:population_heuristics}

@@

By choosing $\theta/N \rightarrow 0$, the second order term in the generator 
will vanish and we expect a deterministic limit,
for which $d/dt \langle f, \eta_t \rangle$ is equal to~\eqref{limit of mean measure equation}.
In other words, the limit is a weak solution to the deterministic equation
\begin{equation}
\label{deterministic limit}
\frac{d\varrho(x)}{dt}
=
    r(x,\varrho(x))
    \DG\big(
        \gamma(\cdot,\varrho) \varrho(\cdot)
    \big)(x)
    + F(x, \varrho(x)) \varrho(x) 
\end{equation}
in the sense of Definition~\ref{defn:weak_solutions},
where $\varrho_t$ is the density of $\eta_t$, if it has a density.

@@

% % % % % % % % % % % % % % % % % %
\subsection{Lineages}

@@

% % % % % % % % % % % % % % % %
\subsection{Definition and generator of the lookdown process}
\label{sec:lookdown_defn}

@@

@@

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Construction tracking individual lines of descent} \label{sec: individual lines of descent}

@@

\subsubsection{Evolution of levels along a whole line of descent}
We now write down expressions
that give the spatial and level dynamic
of a whole line of descent.
In particular, our description is consistent
for any $N \in \mathbb{N}$.

@@

\begin{proposition}[Line of descent construction]
Let $\beta(x, \eta) = \meanq + 2 \covq \grad \log r(x, \eta)$
and $K$ such that $\covq = K K^T$.
Associate with each label $a \in \labelspace$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
as before with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta_0$ and Lebesgue measure.
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Suppose that the spatial locations and level of each line of descent $a$
behave, for $\tau_a \le t < \sigma_a$ as, as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$, with $t < \sigma_a - \tau_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau_a + t$,
location $X_{a \concat j}(\tau_a + t) = X_a(\tau_a + t)$, and level
$$
    U_{a \concat j}(\tau_a + t) = U_a(\tau_a + t)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau_a + t), \eta_{\tau_a + t}) r(X_a(\tau_a + t), \eta_{\tau_a + t}) } .
$$
For each $a$, let $\sigma_a = \inf\{t \ge 0: U_a(t) < \infty\}$.
Then, the processes defined by
\begin{align*}
    \eta_t = \lim_{N \to \infty} \frac{1}{N} \sum_{a : \tau_a \le t < \sigma_a} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
agree in distribution with those of Definitions~\ref{defn:mgale_construction} and~\ref{defn:lookdown_mgale},
respectively.
\end{proposition}

@@

\comment{
    TODO: write overview.
    To prove convergence, we'll be proving
    (a) tightness
    (b) convergence of generators, and
    (c) uniqueness of the limit.

    Then, to get convergence of the lookdown
    we use the population process (somehow).
}

\begin{proof}[Proof of Theorem~\ref{thm:nonlocal_convergence}.]

\comment{INSERT HERE
    narrative of tightness,
    proved in Section \ref{sec:population_tightness_proofs},
    and convergence of generators,
    proved in Section \ref{sec:population_generators_proofs}.
}

% % % % % % % % % % % % %
\subsection{Tightness of empirical measures}
    \label{sec:population_tightness_proofs}
To prove tightness of our population process, 
we first show that the expected total mass of our population process at some fixed time $T>0$,
is controlled uniformly across $\theta$.
This is equivalent to the compact containment condition.

@@

% % % % % % % % % % % % %
\subsection{Convergence of generators}
    \label{sec:population_generators_proofs}

@@

\begin{lemma}[Convergence of generators for the population process]
As $N \to \infty$,
$(\eta^{N}_t)_{t \geq 0}$ converges in $\mathcal{D}_{[0,\infty)}(\mathcal{M}_f(\mathbb{R}^d))$
to $(\eta^{N}(t))_{t \geq 0}$.
\end{lemma}
\begin{proof}
We use Theorem 4.8.10 in \cite{EK}.
First of all, the set of functions
$\{F_f(\eta):= F(\langle f, \eta \rangle ),~
F \in \mathcal{C}^{2}(\mathbb{R}), ~
f \in \mathcal{C}_{0}(\mathbb{R}^d)\}$
is a dense and convergence-determining 
subset of functions in $\mathcal{M}_f(\mathbb{R}^d)$.
Therefore, it suffices to show that for any $t>0$,
\begin{equation}
    \label{eq: Convergence Condition}
\lim_{N \to \infty}
\mathbb{E}\left[
\left(
F_f(\eta^{N}_{t+\tau})-F_f(\eta^{N}_t)
-\int_{t}^{t+\tau}\Pgen^{\infty}F_f(\eta^{N}_s)ds
\right)
\prod_{i=1}^{k}\langle h_i,\eta^{N}_{t_i} \rangle
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+\tau$,
and $h_1,...,h_k \in \mathcal{C}_{0}(\mathbb{R}^d)$.

@@

% % % % % % % % % % % % %
\subsection{Uniqueness of the limit}
    \label{sec:uniqueness_proofs}

Here are the proofs of Corollaries~\ref{cor:nonlocal_pde_limits}
and \ref{cor:superprocess_uniqueness},
giving conditions in which the limiting process is unique
(in which case we have convergence, not just on subsequences).

@@

\comment{... for the cases in which we know this.
This then shows that we can take the local, deterministic limit
by taking $\theta, N \to \infty$ first, and then $\epsilon \to 0$ after.
}

@@

\subsubsection{Porous Media Equation}

@@

@@

\comment{
    Convergence of generators along (something like) the diagonal,
    taking $\epsilon \to 0$ at the same time as the other coefficient.
    (We won't prove tightness here.)
}

@@

% % % % % % % % % % % % %
\subsection{Tightness of the Lines of Descent}
    \label{sec:lookdown_tightness_proofs}

>>>>

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

make sure to put in here that we get convergence of the FKPP

Note: use "reproductive value" instead of "long-term fitness"
and say clearly that "lineage" means "ancestral lineage".

@@

\comment{
    Make sure that we discuss the difference to Flandoli's model
    (one of the few others that show convergence of FKPP),
    and say that a novelty of our results is that we show convergence
    for FKPP and broader models.
}

@@

% % % % % % % % % % % % % %
\paragraph{Structure of the paper}
In the paper we study scaling limits of spatial populations,
obtaining convergence of both the population process
(i.e., the population density as a function of time,
although strictly speaking it is a measure that may not have a density)
as well as of lineages traced back through such a population.
We retain linages through the rescaling process
by means of a lookdown construction.
First, in section~\ref{sec: Model and main results},
we describe the model and the main theorems
(Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}).
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results,
namely, X Y and Z \comment{TODO}.
In section \ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them.
In Section \ref{sec:lookdown} we define and discuss the lookdown process that retains the levels.
This is followed by the technical proofs in section \ref{sec:proofs}.

@@

\plr{TODO: check the conditions below (esp on the derivatives of $F$)}

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    r(x, \smooth{r} \eta(x))
    \gamma(\smooth{\gamma} \eta(x))
    - \frac{1}{\theta}
    F(x, \smooth{F} \eta(x)) .
\end{align}
\end{definition}

@@

% % % % % % % % % % % % % % %
\subsection{Scaling limits of the population density process}

Our main results depend on two dichotomies:
Is the limiting process deterministic or a (generalized) superprocess?
And, are interactions local in the limit or not?
Below we have results for deterministic limits with local and nonlocal interactions,
and for superprocess limits with nonlocal interactions.
See Figure \ref{fig:super_vs_det_2d} for illustrative simulations.
We suspect that limits are possible in the remaining (local, superprocess) case
only in $d=1$, but do not discuss the case further.

@@

    The sequence of stochastic processes
    $(\eta^N_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f:\mathbb{R}^d \rightarrow \mathbb{R}$,
    \begin{align} \label{eqn:mgale_problem}
        \begin{split}
        M_t
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \smooth{\gamma} \eta_{s}(x))
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \smooth{r} \eta_{s}(\cdot))
                \right)(x)
                \\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \smooth{F} \eta_{s}(x)),
                \eta_{s}(dx)
            \big\rangle ds,
        \end{split}
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x,  \eta_{s} \right)
            r\left(x, \eta_{s} \right)+\mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\comment{
    TODO: put in what Kurtz \& Xiong have to say about uniqueness,
along the lines of a note saying ``For some conditions under which the limit is known to be unique,
see Kurtz \& Xiong and/or Dawson \& Li.''
}

@@

% % % % % % % % % % % %
\subsection{Ancestral lineages in the scaling limit}

@@

% % % % % % % % % % % %
\subsection{Beyond reaction-diffusion equations: a population described by the porous medium equation}
\comment{look up what the technical term is for the nonlinear diffusive bit?
    and, does "reaction-diffusion" mean parabolic?}

@@

% % % % % % % % % % % %
\subsection{Different types of traveling waves}

@@

% % % % % % % % % % % % % % % %
\subsection{Lineage motion is not uniquely determined by population density}

@@

% % % % % % % % % % % % % % % % % %
\subsection{The population density}
    \label{sec:population_heuristics}

@@

By choosing $\theta/N \rightarrow 0$, the second order term in the generator 
will vanish and we expect a deterministic limit,
for which $d/dt \langle f, \eta_t \rangle$ is equal to~\eqref{limit of mean measure equation}.
In other words, the limit is a weak solution to the deterministic equation
\begin{equation}
\label{deterministic limit}
\frac{d\varrho(x)}{dt}
=
    r(x,\varrho)
    \DG\big(
        \gamma(\cdot,\varrho) \varrho(\cdot)
    \big)(x)
    + F(x, \varrho) \varrho(x) 
\end{equation}
in the sense of Definition~\ref{defn:weak_solutions},
where $\varrho_t$ is the density of $\eta_t$, if it has a density.

\comment{
    Point out that this is still with $\epsilon > 0$,
    and put in heuristics for why $N \epsilon^d / \theta \to \infty$
    and $\theta \epsilon^2 \to \infty$
    are plausible conditions for the diagonal convergence.
    (Because we want (a) a macroscopic number of neighbors relative to $\theta$,
    and (b) to make the initial condition smooth enough
    since we start with an $\epsilon^2$-smoothed initial condition,
    and we want this to look like a bunch of $1/\theta$-jumps have already happened.
}

@@

% % % % % % % % % % % % % % % % % %
\subsection{Motion of ancestral lineages}

@@

% % % % % % % % % % % % % % % %
\subsection{Lookdown construction of the population density process}
\label{sec:lookdown_defn}

@@

Now that the lookdown process has been defined,
we can present the main convergence theorem that is analogous
to Theorem~\ref{thm:nonlocal_convergence} for the population process.

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ be as defined in Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    The sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$ which satisfies,
    for every $f: \IR^d \to \IR$
    \plr{TODO: write down limiting martingale characterization}
\end{theorem}

@@

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Explicit construction of lines of descent}
    \label{sec: individual lines of descent}

@@

\comment{
    Note here that although we have a single construction that couples
    the processes across all $N$,
    we don't necessarily have pointwise a.s.\ convergence of trajectories of the process.
    Furthermore, individuals at all levels may affect others through the population process,
    and in the deterministic case arbitrarily high levels may drift downwards.
    HOWEVER, this does suggest that a good approximation to the genealogies
    might be obtained by simulating up until a sufficiently high level
    that we have a good approximation to the population process.
}

% % % % % % % % % % % % %  % % %
\subsubsection{Lines of descent in the scaling limit}

We now write down expressions
that give the spatial and level dynamic
of a whole line of descent.
In particular, our description is consistent
for any $N \in \mathbb{N}$.

@@

\begin{proposition}[Line of descent construction]
Let $\beta(x, \eta) = \meanq + 2 \covq \grad \log r(x, \eta)$
and $K$ such that $\covq = K K^T$.
Associate with each label $a \in \labelspace$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
as before with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta_0$ and Lebesgue measure.
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Suppose that the spatial locations and level of each line of descent $a$
behave, for $\tau_a \le t < \sigma_a$ as, as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$, with $t < \sigma_a - \tau_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau_a + t$,
location $X_{a \concat j}(\tau_a + t) = X_a(\tau_a + t)$, and level
$$
    U_{a \concat j}(\tau_a + t) = U_a(\tau_a + t)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau_a + t), \eta_{\tau_a + t}) r(X_a(\tau_a + t), \eta_{\tau_a + t}) } .
$$
For each $a$, let $\sigma_a = \inf\{t \ge 0: U_a(t) < \infty\}$.
Then, the processes defined by
\begin{align*}
    \eta_t
    =
    \lim_{u_0 \to \infty} \frac{1}{u_0}
        \sum_{a : \tau_a \le t < \sigma_a \;\&\; U_a(t) < u_0} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
agree in distribution with those of Definitions~\ref{defn:mgale_construction}
and~\ref{defn:lookdown_mgale}, respectively.
\end{proposition}

@@

\comment{TODO structure of this section}

% % % % % % % % % % % % %
\subsection{Convergence of the population density process}
    \label{sec:population_density_proof}

In this section we prove the main theorem on convergence
of the population density process, Theorem~\ref{thm:nonlocal_convergence}.
This is usually implied by convergence of the lookdown process
(see \citet{kurtz/rodrigues:2011} and \citet{KE});
however in our situation because the parameters depend on the population process
we will use convergence of the population process
in the proofs of tightness and convergence for the lookdown process.

\begin{proof}[Proof of Theorem~\ref{thm:nonlocal_convergence}.]
Here, we need to prove that for any sequence $N_k \to \infty$,
there is a subsequence $N_{k(j)}$ along which $(\eta^{N_{k(j)}}_t)_{t\ge 0}$ converges in distribution
(i.e., tightness for the family of distributions on c\`adl\`ag measure-valued paths),
and that any limit is a solution to the martingale problem stated in the Theorem.
This would convergence along $N_k$ to a unique limit
if the solution to the martingale problem was unique;
however, we only know this to be the case in certain situations.

We prove tightness of $\eta^N$ as $\mathcal{D}_{\cmeasures}([0,\infty))$-valued processes,
where $\cmeasures$ is the space of finite measures on the one-point compactification of $\IR^d$.
To do this, we first prove that $\sup_{0 \le t \le T} \langle 1, \eta^N_t \rangle$ is tight
(a compact containment condition, Lemma~\ref{lem: Compact Containment}).
(The reason we use $\cmeasures$ instead of $\measures$
is so that the set of measures with total mass less than $K$ is compact.)
Next, in Lemma~ we show that the one-dimensional distributions,
$\langle f, \eta^N_t \rangle$ are tight
for each $f \in \mathcal{C}^2_0(\overline{\IR^d})$.
Since this is a dense subset of bounded continuous functions on $\cmeasures$,
together these results allow us to apply Theorem~3.9.1 of \citet{EK}.

Next, we prove that the limit solves the martingale properties
of equations~\eqref{eqn:mgale_problem} and~\eqref{eqn:mgale_variation}
in Lemma~\ref{lem:limit_mgale}
basically by checking the generators converge,
which was mostly done in Section~\ref{sec:population_heuristics}.
\end{proof}

% % % % % % % % % % % % %
\subsubsection{Tightness of empirical measures}
    \label{sec:population_tightness_proofs}
To prove tightness of our population process, 
we first show that the expected total mass of our population process at some fixed time $T>0$,
is controlled uniformly across $\theta$.
This is equivalent to the compact containment condition.

@@

% % % % % % % % % % % % %
\subsubsection{Convergence of generators}
    \label{sec:population_generators_proofs}

@@

\begin{lemma}[Identifying the limit as a solution to the martingale problem]
    \label{lem:limit_mgale}
    \comment{
        Here we need to check that the martingale property is preserved in the limit.
        But this maybe can be shorter?}
As $N \to \infty$,
$(\eta^{N}_t)_{t \geq 0}$ converges in $\mathcal{D}_{[0,\infty)}(\mathcal{M}_f(\mathbb{R}^d))$
to $(\eta^{N}(t))_{t \geq 0}$.
\end{lemma}
\begin{proof}
We use Theorem 4.8.10 in \cite{EK}.
First of all, the set of functions
$\{F_f(\eta):= F(\langle f, \eta \rangle ),~
F \in \mathcal{C}^{2}(\mathbb{R}), ~
f \in \mathcal{C}_{0}(\mathbb{R}^d)\}$
is a dense and convergence-determining 
subset of functions in $\mathcal{M}_f(\mathbb{R}^d)$.
Therefore, it suffices to show that for any $t>0$,
\begin{equation}
    \label{eq: Convergence Condition}
\lim_{N \to \infty}
\mathbb{E}\left[
\left(
F_f(\eta^{N}_{t+\tau})-F_f(\eta^{N}_t)
-\int_{t}^{t+\tau}\Pgen^{\infty}F_f(\eta^{N}_s)ds
\right)
\prod_{i=1}^{k}\langle h_i,\eta^{N}_{t_i} \rangle
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+\tau$,
and $h_1,...,h_k \in \mathcal{C}_{0}(\mathbb{R}^d)$.

@@

@@

\comment{... for the cases in which we know this: FKPP-type and PME.
This then shows that we can take the local, deterministic limit
by taking $\theta, N \to \infty$ first, and then $\epsilon \to 0$ after.
}

@@

\subsubsection{Fisher-KPP-type}

\comment{Insert TwoStepConvergenceV3 with Alison's markups incorporated here.}

\subsubsection{Porous Medium Equation}

@@

\comment{rephrase/shorten this:}

@@

\comment{
    Convergence of generators along (something like) the diagonal,
    taking $\epsilon \to 0$ at the same time as the other coefficient.
    (We won't prove tightness here.)
    This is where OneStepConvergence goes in.
}

@@

% % % % % % % % % % % % %
\subsection{Tightness of the Lines of Descent}
    \label{sec:lookdown_convergence}

Now we prove the main theorem on convergence of the lookdown process, 
Theorem~\ref{thm:lookdown_convergence}.
This follows a similar pattern as the the proofs of convergence for the population process
in Section~\ref{sec:population_density_proof},
although we will need additional estimates to ensure that
trajectories of individual lines of descent converge,
not just the population process.

\begin{proof}[Proof of Theorem~\ref{thm:lookdown_convergence}]
    As in Section~\ref{sec:population_density_proof},
    the theorem will follow from tightness,
    ensuring convergence along subsequences,
    and characterization of the limit.
    This time, the processes $\lp^N$ take values in $\lpmeasures$,
    the space of locally finite measures on space $\times$ levels.
    (They will in fact be point measures, including the limit,
    but that is a consequence of this theorem.)
    Again, tightness will follow from a compact containment condition,
    tightness of one-dimensional distributions,
    and an application of \citet{EK} Theorem~3.9.1.
    We show that the martingale properties are preserved in the limit
    in Section~\ref{sec:lookdown_generator_proofs}.
    \comment{TODO: write and refer to lemma?}

    Lines of descent can escape to infinite level in finite time,
    and so we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    i.e., the topology induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $f(x) g(u)$, where $f \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $g \in C_c([0,\infty))$ is continuous with compact support.
    In this topology, compact containment follows from
    tightness of the spatial and level motions of a single line of descent
    (Lemmas~\ref{lem: tightness of individual spatial trajectory}
    and~\ref{lem:tightness of levels}, respectively),
    and
    tightness of $\lp^N(\IR^d \times [0, u_0])$ for each $u_0$
    (i.e., the number of individuals with level below $u_0$).

    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    However, by Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    This implies that any limit of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.

\comment{MAKE THIS BELOW A LEMMA?}

% BEGIN OLD STUFF
To prove that the lookdown representation is tight, we generalise the vague
    topology to the setting of measure-valued processes. For a detailed
    description of the vague topology, refer to Appendix \ref{sec: Empirical
    Measure Topology}.

Then, we apply Theorem A 2.4 in \cite{kallenberg1997foundations} (Theorem
    \ref{teo: Vague Relative Compactness through Projection}) and use a
    corollary of Stone-Weierstrass Theorem (Lemma \ref{lem: Dense set for test
    functions on empirical measures}) to show that our lookdown representation
    $\{\xi^N_t: N \in \mathbb{N}\}$ is tight as a $\mathcal{M}(\mathbb{R^d}
    \times [0, \infty])$-valued stochastic process in the vague topology as
    long as $\{F_g(\xi^N_t): N \in \mathbb{N}\}$ is tight as real-valued
    stochastic processes for a huge class of $F_g$. Explicitly, $F_g(\xi):=
    \prod_{(x_i,u_i)\in \xi}g(x_i,u_i)$, where $g(x,u)$ satisfies the property
    that $g(x,u) = 1 $ for all $u \geq u_{*}$ for some $u_{*}>0$.
% END OLD STUFF
